---
title: "Week4-Project1"
author: "Simon63"
date: "February 24, 2018"
output: html_document
params:
  csv_file: enter the path for the output file.csv
---

###**NOTE**: Choose Knit with Parameters...

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Libraries
```{r libraries, warning=FALSE, message=FALSE}
library(RCurl)
library(magrittr)
library(stringr)
```

###Load lines from the source text file
```{r load_data}
elo <- readLines('https://raw.githubusercontent.com/simon63/Project1/master/tournamentinfo.txt', warn = F);
head(elo, 10);
```

###Remove 3 lines of the Header
```{r rem_header}
elo <- elo[-c(1:3)]
head(elo, n = 10)
```

###Remove Row-Separator lines
```{r rem_separators}
elo <- subset(elo, str_detect(elo, '^[^-+$]'));
head(elo)
```

###Divide lines into two sets of rows
#####separating odd and even rows into two sets
```{r divide_data}
row1 <- elo[(c(1:length(elo)) %% 2) != 0]
row2 <- elo[(c(1:length(elo)) %% 2) == 0]
message('first set of rows')
head(row1)
message('second set of rows')
head(row2)
```

###Read both sets of rows into two data frame objects
#####and drop the last empty column from the data frames
```{r create_dataframes}
df1 <- read.csv(sep = '|', header = F, text = row1, stringsAsFactors = F)
df2 <- read.csv(sep = '|', header = F, text = row2, stringsAsFactors = F)
# drop the last empty column
df1 <- df1[-11]
df2 <- df2[-11]
knitr::kable(
  head(df1),
  caption = "first data frame"
)
knitr::kable(
  head(df2),
  caption = "second data frame"
)
```

###Extract Opponents' ID values {V4-V10 columns of Data.Frame#1} and convert to numeric
```{r get_ids}
df1$V4 <- df1$V4 %>% str_sub(2, -1) %>% str_trim(side = "both") %>% type.convert
df1$V5 <- df1$V5 %>% str_sub(2, -1) %>% str_trim(side = "both") %>% type.convert
df1$V6 <- df1$V6 %>% str_sub(2, -1) %>% str_trim(side = "both") %>% type.convert
df1$V7 <- df1$V7 %>% str_sub(2, -1) %>% str_trim(side = "both") %>% type.convert
df1$V8 <- df1$V8 %>% str_sub(2, -1) %>% str_trim(side = "both") %>% type.convert
df1$V9 <- df1$V9 %>% str_sub(2, -1) %>% str_trim(side = "both") %>% type.convert
df1$V10 <- df1$V10 %>% str_sub(2, -1) %>% str_trim(side = "both") %>% type.convert
message("data.frame#1 with opponents's ID")
head(df1)
```

###Create a function to calculate *"Average of Pre-Chess Rating of Opponents"*
The function takes in, as parameters, a vector of IDs and the two data frames.
For each ID it picks up the corresponding opponents' IDs (from data.frame#1 omitting NAs)
and looks up the Pre-Chess Rating column (V2) from the corresponding data rows in data.frame#2.
It then creates a pipeline to extract the ratings, convert them to numerics and find the average value
```{r avg_func}

f_avg <- function (v, d1, d2) {
  result <- vector(mode = "double", length = length(v))
  for (i in 1:length(result)) {
    result[i] <- na.omit(d2[unlist(d1[v[i], c(4:10)]),]$V2) %>%
      str_extract_all('R:\\s+\\d+') %>% 
        str_extract('\\d+') %>%
          type.convert %>% 
            mean() %>% 
              round()
  }
  return(result)
}

message("Function test result for Player #1:")
f_avg(c(1), df1, df2)
message("Pre-Chess Rating of Opponents for Player #1:")
na.omit(df2[unlist(df1[1, c(4:10)]), 2]) %>% str_view_all('R:\\s+\\d+')

message("Function test result for Player #37:")
f_avg(c(37), df1, df2)
message("Pre-Chess Rating of Opponents for Player #37:")
na.omit(df2[unlist(df1[37, c(4:10)]), 2]) %>% str_view_all('R:\\s+\\d+')
```

###Build the final data.frame object which will be used to generate a .CSV file
The .CSV file (that could for example be imported into a SQL database) should have the following information for all of the players:  
* Player's Name  
* Player's State  
* Total Number of Points  
* Player's Pre-Rating  
* Average Pre Chess Rating of Opponents  
```{r final_dataframe}
dfResult = data.frame(
  plyr_name = df1$V2 %>% str_trim(side = "both"),
  plyr_state = df2$V1 %>% str_trim(side = "both"),
  plyr_total_pts = df1$V3,
  plyr_pre_rating = df2$V2 %>% str_extract_all('R:\\s+\\d+') %>% str_extract('\\d+') %>% type.convert,
  plyr_oppt_avg_pre_rating = f_avg(c(1:dim(df1)[[1]]), df1, df2)
)
knitr::kable(
  head(dfResult),
  caption = "Resulting Data Frame"
)
```

###Write out the final required .CSV ouput file for this project
```{r csv_file}
write.csv(dfResult, file = params$csv_file)
```